---
title: "Untitled"
output: html_document
date: '2022-07-21'
---

```{r}
loomPath <- system.file(package="SCENIC", "examples/mouseBrain_toy.loom")
library(SCopeLoomR)
loom <- open_loom(loomPath)
exprMat <- get_dgem(loom)
cellInfo <- get_cell_annotation(loom)
close_loom(loom)
```

```{r}
exprMat[1:5,1:5]
```
```{r}
head(cellInfo)
```



```{r}
### Initialize settings
library(SCENIC)
scenicOptions <- initializeScenic(org="mgi", dbDir="/Users/farhadzare/Desktop/Thesis for master/Untitled Folder/databases", nCores=10)

saveRDS(scenicOptions, file="int/scenicOptions.Rds")
```
```{r}
scenicOptions
```

```{r}
### Co-expression network
genesKept <- geneFiltering(exprMat, scenicOptions)
exprMat_filtered <- exprMat[genesKept, ]

#GENIE3/GRNBoost can detect both positive and negative associations. In order to distinguish potential activation from repression, we will split the targets into positive- and negative-correlated targets (i.e. Spearman correlation between the TF and the potential target).


runCorrelation(exprMat_filtered, scenicOptions)
exprMat_filtered_log <- log2(exprMat_filtered+1) 
runGenie3(exprMat_filtered_log, scenicOptions)
```
To build the gene regulatory network: 1. Get co-expression modules 2. Get regulons (with RcisTarget): TF motif analysis) Identify cell states: 3. Score GRN (regulons) in the cells (with AUCell) 4. Cluster cells according to the GRN activity

```{r}
### Build and score the GRN
exprMat_log <- log2(exprMat+1)
scenicOptions@settings$dbs <- scenicOptions@settings$dbs["10kb"] # Toy run settings
scenicOptions <- runSCENIC_1_coexNetwork2modules(scenicOptions)
scenicOptions <- runSCENIC_2_createRegulons(scenicOptions, coexMethod=c("top5perTarget")) # Toy run settings
scenicOptions <- runSCENIC_3_scoreCells(scenicOptions, exprMat_log)
```
```{r}
# Optional: Binarize activity
 aucellApp <- plotTsne_AUCellApp(scenicOptions, exprMat_log)
 savedSelections <- shiny::runApp(aucellApp)
```


```{r}
# Save the modified thresholds:
newThresholds <- savedSelections$thresholds
 scenicOptions@fileNames$int["aucell_thresholds",1] <- "int/newThresholds.Rds"
 saveRDS(newThresholds, file=getIntName(scenicOptions, "aucell_thresholds"))
scenicOptions <- runSCENIC_4_aucell_binarize(scenicOptions)

tsneAUC(scenicOptions, aucType="AUC") # choose settings
```
```{r}
# Export:
 saveRDS(cellInfo, file=getDatasetInfo(scenicOptions, "cellInfo")) # Temporary, to add to loom
#export2loom(scenicOptions, exprMat)
```


### Exploring output 
# Check files in folder 'output'
# Browse the output .loom file @ http://scope.aertslab.org
```{r}
nPcs <- c(5) # For toy dataset
 nPcs <- c(5,15,50)
```

The cells can be grouped/clustered based on the regulon activity, either continuous or binarized

```{r}
scenicOptions@settings$seed <- 123 # same seed for all of them
# Run t-SNE with different settings:
fileNames <- tsneAUC(scenicOptions, aucType="AUC", nPcs=nPcs, perpl=c(5,15,50))
fileNames <- tsneAUC(scenicOptions, aucType="AUC", nPcs=nPcs, perpl=c(5,15,50), onlyHighConf=TRUE, filePrefix="int/tSNE_oHC")
# Plot as pdf (individual files in int/):
fileNames <- paste0("int/",grep(".Rds", grep("tSNE_", list.files("int"), value=T), value=T))
```

```{r}
par(mfrow=c(length(nPcs), 3))
fileNames <- paste0("int/",grep(".Rds", grep("tSNE_AUC", list.files("int"), value=T, perl = T), value=T))
plotTsne_compareSettings(fileNames, scenicOptions, showLegend=FALSE, varName="CellType", cex=.5)
```
```{r}
# Using only "high-confidence" regulons (normally similar)
par(mfrow=c(3,3))
fileNames <- paste0("int/",grep(".Rds", grep("tSNE_oHC_AUC", list.files("int"), value=T, perl = T), value=T))
plotTsne_compareSettings(fileNames, scenicOptions, showLegend=FALSE, varName="CellType", cex=.5)
```
```{r}
scenicOptions@settings$defaultTsne$aucType <- "AUC"
scenicOptions@settings$defaultTsne$dims <- 5
scenicOptions@settings$defaultTsne$perpl <- 15
saveRDS(scenicOptions, file="int/scenicOptions.Rds")
```


```{r}
library(SCopeLoomR)
scenicLoomPath <- getOutName(scenicOptions, "loomFile")
loom <- open_loom(scenicLoomPath)

# Read information from loom file:
regulons_incidMat <- get_regulons(loom)
regulons <- regulonsToGeneLists(regulons_incidMat)
regulonsAUC <- get_regulons_AUC(loom)
regulonsAucThresholds <- get_regulon_thresholds(loom)
embeddings <- get_embeddings(loom)
```


```{r}
exprMat_log <- exprMat # Better if it is logged/normalized
aucellApp <- plotTsne_AUCellApp(scenicOptions, exprMat_log) # default t-SNE
savedSelections <- shiny::runApp(aucellApp)
```

```{r}
print(tsneFileName(scenicOptions))
```


```{r}
tSNE_scenic <- readRDS(tsneFileName(scenicOptions))
aucell_regulonAUC <- loadInt(scenicOptions, "aucell_regulonAUC")

# Show TF expression:
par(mfrow=c(2,3))
AUCell::AUCell_plotTSNE(tSNE_scenic$Y, exprMat, aucell_regulonAUC[onlyNonDuplicatedExtended(rownames(aucell_regulonAUC))[c("Dlx5", "Sox10", "Sox9","Irf1", "Stat6")],], plots="Expression")
```
```{r}
# Save AUC as PDF:
Cairo::CairoPDF("output/Step4_BinaryRegulonActivity_tSNE_colByAUC.pdf", width=20, height=15)
par(mfrow=c(4,6))
AUCell::AUCell_plotTSNE(tSNE_scenic$Y, cellsAUC=aucell_regulonAUC, plots="AUC")
dev.off()
```

```{r}
library(KernSmooth)
library(RColorBrewer)
dens2d <- bkde2D(tSNE_scenic$Y, 1)$fhat
image(dens2d, col=brewer.pal(9, "YlOrBr"), axes=FALSE)
contour(dens2d, add=TRUE, nlevels=5, drawlabels=FALSE)
```

```{r}
regulons <- loadInt(scenicOptions, "regulons")
regulons[c("Dlx5", "Irf1")]
```
```{r}
regulons <- loadInt(scenicOptions, "aucell_regulons")
cbind(onlyNonDuplicatedExtended(names(regulons)))
```
```{r}
regulonTargetsInfo <- loadInt(scenicOptions, "regulonTargetsInfo")
tableSubset <- regulonTargetsInfo[TF=="Stat6" & highConfAnnot==TRUE]
viewMotifs(tableSubset, options=list(pageLength=5))
```
```{r}
motifEnrichment_selfMotifs_wGenes <- loadInt(scenicOptions, "motifEnrichment_selfMotifs_wGenes")
tableSubset <- motifEnrichment_selfMotifs_wGenes[highlightedTFs=="Dlx5"]
viewMotifs(tableSubset) 
```
```{r}
regulonAUC <- loadInt(scenicOptions, "aucell_regulonAUC")
regulonAUC <- regulonAUC[onlyNonDuplicatedExtended(rownames(regulonAUC)),]
regulonActivity_byCellType <- sapply(split(rownames(cellInfo), cellInfo$CellType),
                                     function(cells) rowMeans(getAUC(regulonAUC)[,cells]))
regulonActivity_byCellType_Scaled <- t(scale(t(regulonActivity_byCellType), center = T, scale=T))

ComplexHeatmap::Heatmap(regulonActivity_byCellType_Scaled, name="Regulon activity")
```
```{r}
topRegulators <- reshape2::melt(regulonActivity_byCellType_Scaled)
colnames(topRegulators) <- c("Regulon", "CellType", "RelativeActivity")
topRegulators <- topRegulators[which(topRegulators$RelativeActivity>0),]
viewTable(topRegulators)
```
```{r}
minPerc <- .7
binaryRegulonActivity <- loadInt(scenicOptions, "aucell_binary_nonDupl")
cellInfo_binarizedCells <- cellInfo[which(rownames(cellInfo)%in% colnames(binaryRegulonActivity)),, drop=FALSE]
regulonActivity_byCellType_Binarized <- sapply(split(rownames(cellInfo_binarizedCells), cellInfo_binarizedCells$CellType), 
                                               function(cells) rowMeans(binaryRegulonActivity[,cells, drop=FALSE]))
binaryActPerc_subset <- regulonActivity_byCellType_Binarized[which(rowSums(regulonActivity_byCellType_Binarized>minPerc)>0),]
ComplexHeatmap::Heatmap(binaryActPerc_subset, name="Regulon activity (%)", col = c("white","pink","red"))
```
```{r}
topRegulators <- reshape2::melt(regulonActivity_byCellType_Binarized)
colnames(topRegulators) <- c("Regulon", "CellType", "RelativeActivity")
topRegulators <- topRegulators[which(topRegulators$RelativeActivity>minPerc),]
viewTable(topRegulators)
```

```{r}

# Cell-type specific regulators (RSS): 
regulonAUC <- loadInt(scenicOptions, "aucell_regulonAUC")
rss <- calcRSS(AUC=getAUC(regulonAUC), cellAnnotation=cellInfo[colnames(regulonAUC), "CellType"], )
rssPlot <- plotRSS(rss)
plotly::ggplotly(rssPlot$plot)
```

